---
title: 'Time Series Analysis'
date: '`r Sys.Date()`'
output:
  html_document:
    number_sections: true
    toc: true
    fig_width: 7
    fig_height: 4.5
    theme: cosmo
    highlight: tango
    code_folding: hide
---

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo=TRUE, error=FALSE)
```

## Introduction
This notebook provides a step-by-step guide for fitting an ARIMA model on the stock data, using R.

References:
1. https://a-little-book-of-r-for-time-series.readthedocs.io/en/latest/src/timeseries.html
2. https://sites.google.com/site/econometricsacademy/econometrics-models/time-series-arima-models

## Load libraries 
```{r}
packages =  c("ggplot2", "dplyr", "tidyr", "data.table", 'corrplot', 'gridExtra', 'forecast', 'tseries', 'TSA', 'tibble', 'TTR')

my.install <- function(pkg, ...){
  if (!(pkg %in% installed.packages()[,1])) {
    install.packages(pkg)
  }
  return (library(pkg, ...))
}

purrr::walk(packages, my.install, character.only = TRUE, warn.conflicts = FALSE)
```

## Read Data
```{r}
s_data <- read.csv(file ="../input/all_stocks_5yr.csv")
```

## Data overview
We will use 'summary' and 'str' fuctions of R to look at the data.
```{r}
summary(s_data)
str(s_data)
```

The data has some missing values, which we will replace with zero. Also, 'Date' feature is listed as factor, we will convert that to 'Date' structure.

## Data cleaning
```{r}
s_data[is.na(s_data)] <- 0
s_data$Date <- as.Date(s_data$Date, format = "%Y-%m-%d")
summary(s_data)
str(s_data)
```

## Lets look at some univariate distributions - AllStocks Data
```{r}
options(repr.plot.width=12, repr.plot.height=12) 

p1 = ggplot(s_data, aes(Open)) + geom_histogram(bins = 50, aes(y = ..density..), col = "red", fill = "red", alpha = 0.3) + geom_density()# + xlim(c(0, 1000))

p2 = ggplot(s_data, aes(High)) + geom_histogram(bins = 50, aes(y = ..density..), col = "red", fill = "red", alpha = 0.3) + geom_density()# + xlim(c(0, 1000))

p3 = ggplot(s_data, aes(Low)) + geom_histogram(bins = 50, aes(y = ..density..), col = "red", fill = "red", alpha = 0.3) + geom_density()# + xlim(c(0, 1000))

p4 = ggplot(s_data, aes(Close)) + geom_histogram(bins = 50, aes(y = ..density..), col = "red", fill = "red", alpha = 0.3) + geom_density()# + xlim(c(0, 1000))

grid.arrange(p1,p2,p3,p4, nrow=2,ncol=2)
```

## Time Series Analysis

Before we start with the time series analysis, lets go through the theory in brief.

What is AutoRegressive or AR model:

Autoregressive (AR) models are models where the value of variable in one period is related to the values in the previous period.
AR(p) is a Autoregressive model with p lags.
  
What is Moving Average or MA model:

Moving average (MA) model accounts for the possibility of a relationship between a variable and the residual from the previous period.
MA(q) is a Moving Average model with q lags.

What is ARMA model:
Autoregressive moving average model combines both p auto regressive terms and q Moving average terms, also called ARMA(p,q)

  
Now lets look at some individual stocks and individual time series (Open, Close, High, Low, Volume)

```{r}
tmp <- filter(s_data, High > 100) 
sample(tmp$Name, 10)
i_stock <- filter(s_data, Name == "PSA")
str(i_stock)
```

## Create time series
We will use tsclean and ts function of R to create a time series. 

tsclean() is a convenient method for outlier removal and inputing missing values

ts() is used to create time-series objects
```{r}
## Create a daily Date object
inds <- seq(as.Date("2012-08-13"), as.Date("2017-08-11"), by = "day")

create_ts <- function(col_idx){
  ## Create a time series object
  i_ts <- as.numeric(i_stock[,col_idx]) %>%
    tsclean(replace.missing = TRUE, lambda = NULL) %>%
    ts(start = c(2012, as.numeric(format(inds[1], "%j"))),
             frequency = 365.5)
   return(i_ts)
}
```


## Plotting Time Series - High

There are 5 time series in the data provided - (High, Low, Open, Close, Volume). We will look at the High values first.

```{r}
i_ts = create_ts(which(colnames(i_stock) == "High"))
plot.ts(i_ts, xlab = "Time", ylab = "High value", main = "Time Series", col = "red")
```

## Stationarity
What is stationary time series?

  A stationary process has a mean and variance that do not change overtime and the process does not have trend.
  
  The above time series does not look stationary.
  
  To confirm that we will use "Dickey-Fuller test" to determine stationarity. 
  
Dickey-Fuller test for variable

```{r}
adf.test(i_stock[,which(colnames(i_stock) == "High")], alternative = "stationary", k = 0)
```


## Decomposing Time Series
Decomposing a time series involves separating the time series into trend and irregular components.

```{r}
i_tscomponents <- decompose(i_ts)
plot(i_tscomponents, col = "red")
```

## Differencing a Time Series
Differencing is a common solution used to stationarize the variable.
We will perform differencing using R function diff.
```{r}
i_tsdiff1 <- diff(i_ts, differences=1)
plot.ts(i_tsdiff1, col = "red")
```

The time series (above) appears to be stationary.

## Selecting a Candidate ARIMA Model
The next step is to select appropriate ARIMA model, which means finding the most appropriate values of p and q for an ARIMA(p,d,q) model. You usually need to examine the correlogram and partial correlogram of the stationary time series for this.
To plot a correlogram and partial correlogram, we can use the acf() and pacf() functions in R, respectively.

```{r}
acf(i_tsdiff1, lag.max=60)             # plot a correlogram
acf(i_tsdiff1, lag.max=60, plot=FALSE) # get the autocorrelation values
```


```{r}
pacf(i_tsdiff1, lag.max=60)             # plot a partial correlogram
pacf(i_tsdiff1, lag.max=60, plot=FALSE) # get the partial autocorrelation values
```

Now, we could compare the sample ACF and PACF to those of various theoretical ARMA models. Use properties of ACF & PACF as a guide to estimate plausible models and select appropriate p, q and d. Alternative to this is discussed next.

## Fitting an ARIMA Model
  R provides a function auto.arima, which returns best ARIMA model according to either AIC, AICc or BIC value. The function conducts a search over possible model within the order constraints provided.
  
```{r}
i_tsarima <- auto.arima(i_ts, max.p = 3, max.q = 3, max.d = 3)
i_tsarima
```

## Forecasting using an ARIMA Model
```{r}
i_tsforecasts <- forecast(i_tsarima, h = 60)
plot(i_tsforecasts, col = "red")
```

```{r}
plot.ts(i_tsforecasts$residuals)            # make time plot of forecast errors
ggplot(data.frame(residuals = i_tsforecasts$residuals), aes(residuals)) + geom_histogram(bins = 50, aes(y = ..density..), col = "red", fill = "red", alpha = 0.3) + geom_density()# make a histogram
```

The forecast errors seem to be normally distributed with mean zero and constant variance, the ARIMA model does seem to provide an adequate predictive model


Here we looked at how to best fit ARIMA model to univariate time series.
Next thing that I'll work on is Multivariate Time Series Forecasting using neural net.












